---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Source packages 
Install them if needed (TODO)
```{r}
library(dplyr)
# library(lubridate)  
```


## Loading and preprocessing the data
Assume data is in the same directory as this script

1. If no csv, unzip first
2. Read csv 
3. Format date

```{r}
if (!any(grepl("activity.csv",list.files()))){
  unzip("activity.zip")
}

df <- read.csv("activity.csv")

```



## What is mean total number of steps taken per day?
Summarize steps per day with the dplyr package. Ignore missing data. 

Average 
1. group_by date 
2. summarize dates using sum()
3. summarize daily means using mean()

```{r}
avg_daily_steps <-  group_by(df, date) %>% 
                    summarize(daily = sum(steps, na.rm=TRUE)) %>%
                    summarize(mean_daily = mean(daily, na.rm = TRUE))
```


## What is the average daily activity pattern?
Create a plot of the average daily activity pattern.

1. group_by date
2. summarize dates using me


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
